New Metrics for Assessing Projection Pursuit Indexes, and Guiding Optimisation Choices

### 2024-11-18 jso.qmd
update the modelling result

```{r}
#| label: fit-logistic
sim_df2 <- sim_df |>   
  mutate(n_jellies = n_jellies/10,
         max_tries = max_tries/10,
         long_time = ifelse(time > 20, 1, 0),
         P_J = ifelse(index == "stringy", 0, P_J)) |> 
  rename(proportion = P_J)

mod1 <- glm(proportion ~ smoothness + squintability + d + long_time + n_jellies + max_tries, data = sim_df2, family = binomial)
```

@@

```{r}
#| label: tbl-smoothness-squintability
#| tbl-cap: Parameters estimated from the Gaussian process (outputscale $\eta$, lengthscale $\ell$, smoothness $\nu$, and nugget $\sigma$) and scaled logistic function ($\theta_1$ to $\theta_4$ and $\varsigma$) for the pipe-finding and sine-wave finding problems.  The columns $\nu$ and $\varsigma$ represent the smoothness and squintability measures respectively. 
dt <- tibble(shape = c(rep("pipe", 4), rep("sine", 8)), smoothness) |>
  left_join(squintability) |> 
  rename(d = n, smooth = smoothness) |> 
  mutate(index = ifelse(index == "dcor2d_2", "dcor", index),
         index = ifelse(index == "loess2d", "loess", index),
         index = ifelse(index == "splines2d", "splines", index),
         variance = sqrt(variance)) 
dt |> 
  knitr::kable(
    digits = 2, format = "latex", 
    col.names = c(
      "shape", "index", "d",
      "$\\eta$", "$\\ell$", "$\\nu$", "$\\sigma$",
      "$\\theta_1$", "$\\theta_2$", "$\\theta_3$", "$\\theta_4$", "$\\varsigma$"), linesep = "",
    booktabs = TRUE, escape = FALSE, row.names = FALSE, align = "|llr|rrrr|rrrrr") |> 
  #column_spec(1, border_left = TRUE) |> 
  #column_spec(c(4, 8, 13), border_right = TRUE) |> 
  kable_styling(font_size = 10) 
```

@@

```{r}
#| label: tbl-mod-output
#| tbl-cap: "Jellyfish success rate relative to index properties and jellyfish hyper-parameters. This is the summary from a logistic regression fit to smoothness, squintability, dimension, running time, number of jellyfish and maximum number of tries. Interestingly, squintability and dimension strongly affect jellyfish optimisation success. The number of jellies marginally affects success, but index smoothness, running longer and increasing the number of tries do not."
#| fig-pos: t
library(gtsummary)
set_gtsummary_theme(theme_gtsummary_journal("jama"))
mod1 |> tbl_regression(exponentiate=TRUE, 
                       pvalue_fun = label_style_pvalue(digits = 3),
                       label = list(smoothness = "Smoothness", 
                                    squintability = "Squintability",
                                    d = "Dimension", 
                                    long_time = "Long runtime", 
                                    n_jellies = "Number of jellyfish", 
                                    max_tries = "Maximum number of tries")) |>
  bold_p(t = 0.1)
```

>>>>

```{r}
#| label: fit-logistic
sim_df2 <- sim_df |>   
  mutate(n_jellies = n_jellies/10,
         max_tries = max_tries/10,
         long_time = ifelse(time > 20, 1, 0)
         ) |> 
  rename(proportion = P_J)

mod1 <- glm(proportion ~ smoothness + squintability + d +  n_jellies + max_tries, data = sim_df2, family = binomial)
```

@@

```{r}
#| label: tbl-smoothness-squintability
#| tbl-cap: Parameters estimated from the Gaussian process (outputscale $\eta$, lengthscale $\ell$, smoothness $\nu$, and nugget $\sigma$) and scaled logistic function ($\theta_1$ to $\theta_4$ and $\varsigma$) for the pipe-finding and sine-wave finding problems.  The columns $\nu$ and $\varsigma$ represent the smoothness and squintability measures respectively. 
dt <- tibble(shape = c(rep("pipe", 4), rep("sine", 9)), smoothness) |>
  left_join(squintability) |> 
  rename(d = n, smooth = smoothness) |> 
  mutate(index = ifelse(index == "dcor2d_2", "dcor", index),
         index = ifelse(index == "loess2d", "loess", index),
         index = ifelse(index == "splines2d", "splines", index),
         variance = sqrt(variance)) 
dt |> 
  knitr::kable(
    digits = 2, format = "latex", 
    col.names = c(
      "shape", "index", "d",
      "$\\eta$", "$\\ell$", "$\\nu$", "$\\sigma$",
      "$\\theta_1$", "$\\theta_2$", "$\\theta_3$", "$\\theta_4$", "$\\varsigma$"), linesep = "",
    booktabs = TRUE, escape = FALSE, row.names = FALSE, align = "|llr|rrrr|rrrrr") |> 
  #column_spec(1, border_left = TRUE) |> 
  #column_spec(c(4, 8, 13), border_right = TRUE) |> 
  kable_styling(font_size = 10) 
```

@@

```{r}
#| label: tbl-mod-output
#| tbl-cap: "Jellyfish success rate relative to index properties and jellyfish hyper-parameters. This is the summary from a logistic regression fit to smoothness, squintability, dimension, running time, number of jellyfish and maximum number of tries. Interestingly, squintability and dimension strongly affect jellyfish optimisation success. The number of jellies marginally affects success, but index smoothness, running longer and increasing the number of tries do not."
#| fig-pos: t
library(gtsummary)
set_gtsummary_theme(theme_gtsummary_journal("jama"))
mod1 |> tbl_regression(exponentiate = TRUE, 
                       pvalue_fun = label_style_pvalue(digits = 3),
                       label = list(smoothness = "Smoothness", 
                                    squintability = "Squintability",
                                    d = "Dimension", 
                                    long_time = "Long runtime", 
                                    n_jellies = "Number of jellyfish", 
                                    max_tries = "Maximum number of tries")) |>
  bold_p(t = 0.1)
```

### 2024-11-18 jso.tex
update the modelling result

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{caption}
\usepackage{orcidlink}
\usepackage{algorithm}
\usepackage{float}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\theoremstyle{plain}
\newtheorem{defn}{\protect\definitionname}
\newtheorem{prop}{\protect\propositionname}
\providecommand{\definitionname}{Definition}
\providecommand{\propositionname}{Proposition}
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\@ifpackageloaded{fontawesome5}{}{\usepackage{fontawesome5}}
\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{acacac}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582ec}
\definecolor{quarto-callout-important-color-frame}{HTML}{d9534f}
\definecolor{quarto-callout-warning-color-frame}{HTML}{f0ad4e}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02b875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{fd7e14}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

@@

\begin{tcolorbox}[enhanced jigsaw, bottomtitle=1mm, colbacktitle=quarto-callout-note-color!10!white, title={Algorithm: Jellyfish Optimizer Pseudo Code}, arc=.35mm, breakable, colback=white, opacityback=0, leftrule=.75mm, toprule=.15mm, rightrule=.15mm, toptitle=1mm, colframe=quarto-callout-note-color-frame, titlerule=0mm, coltitle=black, bottomrule=.15mm, left=2mm, opacitybacktitle=0.6]

@@

To assess JSO's performance across various PP problems, two different
data shapes, pipe and a sine wave, are investigated in 6D and 8D spaces
using six different PP indexes: \texttt{dcor2d\_2}, \texttt{loess2d},
\texttt{MIC}, \texttt{TIC}, \texttt{spline}, and \texttt{stringy}, under
varied JSO hyper-parameters. This results in a total of 52 scenarios,
comprising of 24 computed on the pipe data and 28 on the sine-wave data.
Again, JSO is run 50 times for each scenario to calculate the success
rate for each PP problem.

@@

\begin{tabular}{|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr}
\toprule
shape & index & d & $\eta$ & $\ell$ & $\nu$ & $\sigma$ & $\theta_1$ & $\theta_2$ & $\theta_3$ & $\theta_4$ & $\varsigma$\\
\midrule
pipe & holes & 6 & 0.03 & 0.39 & 3.08 & 0.0 & 1.02 & 0.86 & 3.37 & 0.02 & 3.05\\
pipe & holes & 8 & 0.02 & 0.32 & 2.86 & 0.0 & 1.01 & 0.87 & 3.26 & 0.03 & 2.96\\
pipe & holes & 10 & 0.02 & 0.36 & 2.80 & 0.0 & 1.02 & 0.88 & 3.15 & 0.02 & 2.95\\
pipe & holes & 12 & 0.02 & 0.36 & 2.82 & 0.0 & 1.01 & 0.88 & 3.34 & 0.00 & 3.12\\
sine & MIC & 6 & 0.06 & 0.29 & 3.15 & 0.0 & 0.89 & 0.57 & 1.62 & -0.02 & 1.26\\
sine & MIC & 8 & 0.06 & 0.33 & 3.09 & 0.0 & 0.93 & 0.33 & 1.31 & -0.03 & 0.77\\
sine & TIC & 6 & 0.17 & 0.34 & 3.06 & 0.0 & 0.95 & 0.54 & 1.72 & -0.03 & 1.32\\
sine & TIC & 8 & 0.16 & 0.31 & 3.17 & 0.0 & 0.95 & 0.56 & 1.72 & -0.03 & 1.37\\
sine & dcor & 6 & 0.09 & 0.31 & 3.06 & 0.0 & 0.95 & 1.04 & 2.74 & -0.02 & 2.95\\
sine & loess & 6 & 0.24 & 0.45 & 3.53 & 0.0 & 1.02 & 1.04 & 2.65 & 0.08 & 2.76\\
sine & splines & 6 & 0.10 & 0.32 & 3.11 & 0.0 & 1.01 & 1.05 & 2.73 & -0.01 & 3.12\\
sine & stringy & 6 & 0.00 & 1128.28 & 1.03 & 134.8 & 1.05 & 0.01 & 254.73 & 0.05 & 2.96\\
\bottomrule
\end{tabular}
\endgroup{}

@@

Table~\ref{tbl-mod-output} presents the results from fitting a logistic
regression model using the proportion of success as the response and
smoothness, squintability, dimension, running time, number of jellyfish
and maximum number of tries, as predictors. Long runtime is a binary
variable, where greater than 20 seconds is labelled 1. The fit suggests
that JSO success rate is only affected by squintability, dimension and
number of jellies. Interestingly, the smoothness of the index is not a
problem for the JSO - this is promising! Running the optimisation longer
and allowing more tries also does not affect success rate. An increase
in squintability of one unit increases the success rate by 1235\%. As
dimension increases by one the success rate halves. Increasing the
number of jellies by 10, increases the succes rate by 27\%.

\setlength{\LTpost}{0mm}

@@

\toprule
\textbf{Characteristic} & \textbf{OR} \textbf{(95\% CI)}\textsuperscript{\textit{1}} & \textbf{p-value} \\ 
\midrule\addlinespace[2.5pt]
Smoothness & 89.1 (1.23 to 4,865,159) & 0.289 \\ 
Squintability & 12.4 (3.08 to 125) & \textbf{0.006} \\ 
Dimension & 0.57 (0.30 to 0.96) & \textbf{0.048} \\ 
Long runtime & 0.17 (0.00 to 4.96) & 0.357 \\ 
Number of jellyfish & 1.27 (0.99 to 1.68) & \textbf{0.069} \\ 
Maximum number of tries & 1.10 (0.82 to 1.52) & 0.524 \\ 
\bottomrule

@@

\begin{minipage}{\linewidth}
\textsuperscript{\textit{1}}OR = Odds Ratio, CI = Confidence Interval\\
\end{minipage}

>>>>

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
\usepackage{caption}
\usepackage{anyfontsize}
\usepackage{orcidlink}
\usepackage{algorithm}
\usepackage{float}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\theoremstyle{plain}
\newtheorem{defn}{\protect\definitionname}
\newtheorem{prop}{\protect\propositionname}
\providecommand{\definitionname}{Definition}
\providecommand{\propositionname}{Proposition}
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\@ifpackageloaded{fontawesome5}{}{\usepackage{fontawesome5}}
\definecolor{quarto-callout-color}{HTML}{909090}
\definecolor{quarto-callout-note-color}{HTML}{0758E5}
\definecolor{quarto-callout-important-color}{HTML}{CC1914}
\definecolor{quarto-callout-warning-color}{HTML}{EB9113}
\definecolor{quarto-callout-tip-color}{HTML}{00A047}
\definecolor{quarto-callout-caution-color}{HTML}{FC5300}
\definecolor{quarto-callout-color-frame}{HTML}{acacac}
\definecolor{quarto-callout-note-color-frame}{HTML}{4582ec}
\definecolor{quarto-callout-important-color-frame}{HTML}{d9534f}
\definecolor{quarto-callout-warning-color-frame}{HTML}{f0ad4e}
\definecolor{quarto-callout-tip-color-frame}{HTML}{02b875}
\definecolor{quarto-callout-caution-color-frame}{HTML}{fd7e14}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

@@

\begin{tcolorbox}[enhanced jigsaw, colframe=quarto-callout-note-color-frame, opacityback=0, breakable, toptitle=1mm, opacitybacktitle=0.6, bottomtitle=1mm, colbacktitle=quarto-callout-note-color!10!white, toprule=.15mm, colback=white, title={Algorithm: Jellyfish Optimizer Pseudo Code}, coltitle=black, bottomrule=.15mm, leftrule=.75mm, titlerule=0mm, arc=.35mm, rightrule=.15mm, left=2mm]

@@

To assess JSO's performance across various PP problems, two different
data shapes, pipe and a sine wave, are investigated in 6D and 8D spaces
using six different PP indexes: \texttt{dcor2d\_2}, \texttt{loess2d},
\texttt{MIC}, \texttt{TIC}, \texttt{spline}, and \texttt{stringy}, under
varied JSO hyper-parameters. This results in a total of 54 scenarios,
comprising of 24 computed on the pipe data and 30 on the sine-wave data.
Again, JSO is run 50 times for each scenario to calculate the success
rate for each PP problem.

@@

\begin{tabular}{|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr|llr|rrrr|rrrrr}
\toprule
shape & index & d & $\eta$ & $\ell$ & $\nu$ & $\sigma$ & $\theta_1$ & $\theta_2$ & $\theta_3$ & $\theta_4$ & $\varsigma$\\
\midrule
pipe & holes & 6 & 0.03 & 0.39 & 3.08 & 0.00 & 1.02 & 0.86 & 3.37 & 0.02 & 3.14\\
pipe & holes & 8 & 0.02 & 0.32 & 2.86 & 0.00 & 1.01 & 0.87 & 3.26 & 0.03 & 3.04\\
pipe & holes & 10 & 0.02 & 0.36 & 2.80 & 0.00 & 1.02 & 0.88 & 3.15 & 0.02 & 3.05\\
pipe & holes & 12 & 0.02 & 0.36 & 2.82 & 0.00 & 1.01 & 0.88 & 3.34 & 0.00 & 3.20\\
sine & MIC & 6 & 0.06 & 0.31 & 3.08 & 0.00 & 0.89 & 0.57 & 1.62 & -0.03 & 1.38\\
sine & MIC & 8 & 0.03 & 0.32 & 2.75 & 0.00 & 0.93 & 0.33 & 1.31 & -0.04 & 0.83\\
sine & TIC & 6 & 0.16 & 0.31 & 3.08 & 0.00 & 0.95 & 0.54 & 1.72 & -0.03 & 1.43\\
sine & TIC & 8 & 0.07 & 0.30 & 2.72 & 0.00 & 0.95 & 0.56 & 1.72 & -0.03 & 1.48\\
sine & dcor & 6 & 0.08 & 0.29 & 3.13 & 0.00 & 0.95 & 1.04 & 2.74 & -0.03 & 3.20\\
sine & loess & 6 & 0.24 & 0.46 & 3.56 & 0.00 & 1.02 & 1.04 & 2.65 & 0.07 & 3.02\\
sine & splines & 6 & 0.10 & 0.32 & 3.11 & 0.00 & 1.01 & 1.05 & 2.73 & -0.02 & 3.40\\
sine & stringy2 & 6 & 0.01 & 0.09 & 0.71 & 0.97 & 0.65 & 0.05 & 7.78 & 0.06 & 0.39\\
sine & skinny & 6 & 0.05 & 0.19 & 2.24 & 0.01 & 0.83 & 0.00 & 2.98 & 0.05 & 0.00\\
\bottomrule
\end{tabular}
\endgroup{}

@@

Table~\ref{tbl-mod-output} presents the results from fitting a logistic
regression model using the proportion of success as the response and
smoothness, squintability, dimension, running time, number of jellyfish
and maximum number of tries, as predictors. Long runtime is a binary
variable, where greater than 20 seconds is labelled 1. The fit suggests
that JSO success rate is only affected by squintability, dimension and
number of jellies. Interestingly, the smoothness of the index is not a
problem for the JSO - this is promising! Running the optimisation longer
and allowing more tries also does not affect success rate. An increase
in squintability of one unit increases the success rate by 483\%. As
dimension increases by one the success rate halves. Increasing the
number of jellies by 10, increases the succes rate by 8\%.

\begingroup
\fontsize{12.0pt}{14.4pt}\selectfont
\setlength{\LTpost}{0mm}

@@

\toprule
\textbf{Characteristic} & \textbf{OR} \textbf{(95\% CI)}\textsuperscript{\textit{1}} & \textbf{p-value} \\ 
\midrule\addlinespace[2.5pt]
Smoothness & 0.64 (0.09 to 14.1) & 0.673 \\ 
Squintability & 4.83 (1.99 to 14.2) & {\bfseries 0.001} \\ 
Dimension & 0.57 (0.36 to 0.85) & {\bfseries 0.008} \\ 
Number of jellyfish & 1.23 (0.98 to 1.59) & {\bfseries 0.086} \\ 
Maximum number of tries & 1.08 (0.82 to 1.44) & 0.593 \\ 
\bottomrule

@@

\begin{minipage}{\linewidth}
\textsuperscript{\textit{1}}OR = Odds Ratio, CI = Confidence Interval\\
\end{minipage}
\endgroup
